---
title: "Exam 2016 "
author: "Thando Shabalala"
date: "August 31, 2016"
output: html_document
---
# Question 1

##Null Hypothesis
* There is no relationship between oral body temperature and heart rate

##Alternative Hypothesis
* There is a relationship exists between oral body temperature and heart rate.

##Statistical Test(s)
* ANOVA
## Assumptions of the statistical test
* the distribution is normal
* equal variance accros groups
*data are unmatched
*errors are independent

```{r question1}
library(tidyr)
library(dplyr)
library(ggplot2)
library(knitr)

# View and Import Dataset Question 1 CSV
q1 <- read.csv("question1.csv")
qdset <- tbl_df(q1)
q1

# Plots
plot(q1$male ~ q1$body_temperature,
     col="pink",main = "Mean oral body temperature and heart rate at rest in male and female students",ylab = " Heart Rate",xlab = "Oral Body Temperature")
points(q1$female)
legend(37.5,85, c("Male", "Female"), fill = c("pink", "black"))

```
##Outcome Analysis
* There is a relationship between the body temperature and heart rate especially in the males since they have more data.Therefore we can reject the null hypothesis.

#Question2

##Null Hypothesis
* Intoxication does not affect the handedness and ataxic walking.

##Alternative Hypothesis
* Intoxication affects handedness and ataxic walking.

##Statistical Test(s)

* Wilcoxin rank-sum test

##Assumptions of the statistical test
* The samples are drawn from populations of the same shape
* errors are independent
* Data are unmatched 

```{r question2}
library(tidyr)
library(dplyr)
library(ggplot2)
library(knitr)

# View and Import Dataset Question 2 CSV

q2 <- read.csv("question2.csv")
q2data<- tbl_df(q2)
q2data

# Plots
boxplot(q2data$first_stumble ~ q2data$handedness)

par(mfrow = c(1,2),mar = c(4, 4, 2, 1),oma = c(0, 0, 1, 0)) 

mosaicplot(q2data$handedness ~ q2data$first_stumble, main = "Handedness against position of first stumble", ylab = "First Stumble (1= rightwards, 2 = leftwards)", xlab = "Handedness (1 = right-handed, 2 = left-handed)")

mosaicplot(q2data$final_position ~ q2data$handedness,main = "Handedness against final position",ylab = "Final Stimble (1= rightwards, 2 = leftwards)",xlab = "Handedness (1 = right-handed, 2 = left-handed)")

wilcox.test(first_stumble ~ handedness, data = q2data, paired = FALSE)

wilcox.test(final_position~handedness,data = q2data,paired = FALSE)

par(mfrow = c(1,1),mar = c(4, 4, 2, 1),oma = c(0, 0, 1, 0))            
```
##Outcome Analysis
Intoxication affects how a person is poitioned in their gait, there is a relationship between stumbling and handedness as well as stumbling and positioning.The pvalue was 0.04.So we can reject the null hypothesis.



#Question3

##Null Hypothesis
* Running time doesn't calorie consumption.

##Alternative Hypothesis
* Running time improves calorie consumption.

```{r question3}
library(tidyr)
library(dplyr)
library(ggplot2)
library(knitr)

# View and Import Dataset Question 3 CSV
q3 <- read.csv("question3.csv")
q3cal <- tbl_df(q3)
q3cal

# Plot

plot(q3cal$calories ~ q3cal$time,
      main = "Relationship between an amateur runner's running time and calories consumed", ylab = "Calories Consumed (cal)", xlab = "Running Time (s)")

p <- qplot(x = time, y = calories,data = q3cal,main = "Relationship between an amateur runner's running time and calories consumed",ylab = "Calories Consumed (cal)",xlab = "Running Time (s)")

#Pearson Correlation and print the summary statistics

q3cal.cor <- with(q3cal, cor.test(method = 'pearson', x = time, y = calories))
q3cal.cor

#Annotate r and p values onto the plot

p + annotate("text", x = 2150, y = 420, label = "r = -0.7") + annotate("text", x = 2150, y = 410, label = "p = 0.002")

# View
par(mfrow = c(1,2),mar = c(4, 4, 2, 1),oma = c(0, 0, 1, 0))

#Summary of linear regression
q3cal.reg <- lm(calories ~ time, data = q3cal)
summary(q3cal.reg)

#Plot

plot(x = q3cal.reg$fitted.values,y = q3cal.reg$residuals,xlab = "Fitted Values",ylab = "Residuals",main = "Homoskedasticity of residuals")
abline(h = 0)

# 
qqnorm(q3cal.reg$residuals, main = "Normal residula distribution")
qqline(q3cal.reg$residuals)

# Overall Title
mtext("Regression analysis", outer = TRUE)
```
##Statistical Test(s)
 *Linear regression 
##Assumptions of the statistical test
* Ran a linear regression to see if there was a linear trend between x and y
* A qqplot to check if the residuals were normally distributed 
*homoskedastic test to see if the residuals had the same variance for all fitted values of x and y.

##Intepretation
* The plot shows that there is a strong inverse relationship between running time and calorie consumption.The more time the person runs the more time the more time they spend running the less the more calories they consume.

##Amount of calories consumed 
* At 30 minutes the person would consume about 420 calories

#question4
```{r question4 ,fig.path = './figures/', dev = 'pdf'}

foo <- rnorm(10000, mean = 60, sd = 3) # final mark
bar <- rnorm(10000, mean = 68, sd = 5) # project mark
baz <- rep(seq(from = 1997, to = 2006), each = 1) # years

year <- sample(baz, 150, replace = TRUE,prob = c(0.05, 0.05, 0.08, 0.08, 0.1, 0.13, 0.14, 0.13, 0.12, 0.12))

project_mark <- sample(bar, 150, replace = TRUE)

final_mark <- sample(foo, 150, replace = TRUE)

q4plot <- data_frame(year = year,project_mark = round(project_mark, 1),final_mark = round(final_mark, 1)) %>%
arrange(year)

# View
q4plot

# Tidying Data
data_tidy <- q4plot %>%
    mutate(year = factor(year)) %>%
    gather(key = key, value = value, -year) %>%
    arrange(year)

# Plot
with (data_tidy, boxplot(value ~ key,main = "Main",ylab = "percentage"))
```
